Carbon Nucleosynthesis

### 2023-04-10 thesis/tex/main.tex
more thesis edits

\begin{titlepage}
   \begin{center}
       \vspace*{5\baselineskip}
       \textbf{The Galactic Chemical Evolution of Carbon} \\
       \vspace*{3\baselineskip}
        Undergraduate Research Thesis\\
       \vspace*{3\baselineskip}
    Presented in partial fulfillment of the requirements for graduation \textit{with research distinction in Astronomy and Astrophysics} in the College of Arts and Sciences of The Ohio State University \\
       \vspace*{3\baselineskip}
        by \\
       \vspace*{3\baselineskip}
       {Daniel Alexander Boyea}\\
       \vspace*{3\baselineskip}
       The Ohio State University\\
       April 2023\\
       \vspace*{3\baselineskip}
       Project Advisor: David Weinberg, Department of Astronomy
       % Rolando Vald\'es Aguilar\\
       \vfill
   \end{center}
\end{titlepage}

@@

Here, I aim to understand C enrichment---where it is produced and how its abundances evolve over time. Why study C? C is unique nucleosynthetically, one of few light elements (along with N) produced in Asymptotic Giant Branch (AGB) stars \citep[e.g.][]{jennifer19, KL14}. C and N are well studied elements as they are easy to observe in stellar spectra, even at the lowest metallicities \citep{FN15} SOMETHING ELSE?. Additionally, C and N are used as age indicators in red giant branch stars \citep[e.g.][]{martig16, MG15, fiorenzo+21}.
While we know C is produced in AGB stars and core collapse supernovae (CCSNe), we still have limited understanding of the importaince and metallicity-dependence of each process.

Empirically, gas-phase [C/O] ratios decrease with increasing [O/H] at the lowest metallicitis---in high redshift damped Lyman-alpha systems (DLAs)---and then [C/O] increases above $\rm [O/H] \approx 1$ (\citealt{berg+19,cooke+17}; see discussion in SECTION).

\cite{james+22} examend GCE models of N, an elements whose production is closely related to C. \cite{james+22} found trends in N and O abundances can be explained by the metallicity dependence of relatice N and O yields. Here, I extend their models to C and show that similar constraints on C and O relative yields can be obtained from observed abundance ratios.

We also know that there is evidence for non-monatonic C yield patterns. 
pattern where C is more readily produced by both high metallicity and very low metallicity stars (CEMPS, early GCE results), but our picture. \citep{FN15,cooke+17,berg+19}.

Despite their centrol role in GCE models, nucleosynthetic yields predicted by stellar evolution models are rife with uncertainties. Poorly understood processes, including mass loss, opacity, nuclear reaction rates, rotational mixing, and convection/convective boundaries each shape the production rates of atomic nuclei in stellar interios \citep{KL14,V13, LC18, emily+21}.
Additionally, changes in massive star's explodabinitily can introduce a factor of \~2 change in population averaged-yields \citep{emily+21}.

Stellar evolution models predict stellary yields, playing a crucial role in GCE.  However, these models are rife with uncertainties, limited by our understanding of critical physical processes such as nuclear reaction rates, convection, rotational mixing, opacity, and mass loss CITATIONS. Changes in opacity, rotation, reaction rates, and explodability can substantially affect C yields

GCE models typically use unmodified nucleosythetic predictions and attempt to create models matching observaions by varying star formation history and other parameters. I instead use the observations as a constraint, determining which yield assuptions are consistant with galactic abundance trends.

While there are many stellar measurments of C, accurately determining the birth abundance is more challenging. For example, when a star becomes a RGB star, material from the CNO-processed core is mixed with the envelope in First Dredge Up (FDU) \citep{fiorenzo+21,KL14}. Since this core-processed material is now C depleted and N enhanced, measurments of this star's atmosphere will no longer reflect the birth abundances of the star.  Additionaly, C is faint in HII regions and determining gas-phase abundances is challenging \citet{skillman+20}.
Other surveys, such as GALAH, have unreliable C measurments with low detection rates \citep{emily+19}.

AGB stars are known to be importaint (cite) from stellar model predictions, previous GCE, and evidence from abundance patterns. (are there detections in PN?)

CCSne are also important (.... SN reminant measurments)

To this end, I use a sample of subgiant stars from APOGEE \citep{apogee17}. According to stellar evolution models (CITES), these stars have not yet experienced dredge-up. Their photospheric abundances should thus represent their birth composition.  In Fig. \ref{fig:subgiants}, I plot this sample, selected by criteria in \citet[see Appendix \ref{sec:jack}]{jack_subgiant}. These emperical resolults contain information on the relative C and O yields, which I use later in this work.

@@

Theoretical models of stellar nucleosynthesis provide the starting point of this investigation. I focus on three primary nucleosynthetic pathways: Asymptotic Giant Branch (AGB) stars, Core Collapse Supernovae (CCSNe), and Type Ia Supernovae (SNe Ia). Each process has unique timescales and yields, tracable through the tools of GCE. While C is only produced in AGB stars and CCSNe, comparing C to Fe, which is produced in SNe Ia, adds a valuable constraint on the delay-time distribution (DTD) of C enrichment.

When a Single Stellar Population (SSP) forms, CCSNe are the first enrichment events. CCSNe provide light elements such as C, O, and Mg and heavier elements such as Fe and beyond. Metallicity-independent yields from CCSNe are the only statisticly significant source of O and Mg ($\alpha$ elements). Next, low mass stars begin to reach the end of their lives. The dying breaths of AGB stars are known to be importaint sources of C, N, and neutron capture elements.  Finally, white dwarfs explode, releasing Fe and other iron-peak elements.

Stellar measurments of Mg are more reliable, but O abundances are easier to measure in HII regions. 
I therefore use Mg abundances from the \citet{jack_subgiant} subgiant sample as my primary emperical constraint. However, I use O abundances for gas-sphase measurments and low-metallicity stars (\S \ref{sec:gas}). I model both Mg and O as metallicity-independent population-averaged yields dominated by massive stars. ${\rm [O/Mg]}\approx 0$ across a wide range of metallicities in APOGEE \citep{Weinberg+19, Weinberg+22}. 
I adopt O and Mg relative yields given by the solar mixture of \cite{asplund+09} <MATH>.

To calculate the IMF-weighted\footnote{IMF: Initial Mass Function} net yields of C, I use the Versatile Integrator for Chemical Evolution code (\VICE\footnote{\VICE~is available at \url{https://github.com/giganano/VICE}}).
As stellar evolution models predict per-star yields of a given mass, I need to integrate over the IMF to determine the total yields in terms of star formation.

In the interest of focusing on carbon yields, I hold yields of other elements fixed. 
I adopt yields from \citet{james+21, james+22}.
In Table \ref{tab:fiducial_mod}, I tabulate yield values, in units of a stellar population's birth mass.
Also following \citet{james+21, james+22}, I take the SNeIa Delay Time distribution to be a
$t^{-1.1}$ law suggested by the observations of \citet{maoz+12}.

@@

An AGB star is a low mass ($\lesssim 8 M_{\sun}$) star during its final phase of evolution.  In an AGB star, two competing processes determine the outcome of C production: Third Drege Up (TDU) and Hot Bottom Burning (HBB).  DU accompianies thermal puslations in AGB stars. Durring a TDU event, material from the CO core is mixed with material exposed to partial He-shell burning, increasing C and O abundances in the envelope \citep{KL14}.
HBB is the activation of proton capture reactions, and the CNO cycle%
\footnote{The CNO (I) cycle is
\isotope[12]{C}(p, $\gamma$)
\isotope[13]{N}($\beta^+ \nu_e$)
\isotope[13]{C}(p, $\gamma$)
\isotope[14]{N}(p, $\gamma$)
\isotope[15]{O}($\beta^{+}\nu_e$)
\isotope[15]{N}(p, \alp)
\isotope[12]{C}.
 Other minor branches of the CNO cycle are not relevant here.
 \citep{solar-fusion}.
}
at the base of the convective envelope. Because the $^{14}$N(p, $\alpha$) proton capture is the slowest component of the CNO cycle \citep{solar-fusion}, the CNO cycle converts nearly all \isotope[12]{C} into \isotope[14]{N}.
As a result, when both TDU and HBB occur, \isotope[12]{C} yields are lowered (see discussion in \citealt{james+22} and \citealt{ventura+13}).

In this work, I explore four different sets of AGB star yield tables from literature, providing well sampled grids in metallicity and mass for use in chemical evolution models. 
\begin{itemize}
    \item C11: \citet{cristallo+11} and \citet{cristallo+15}
    \item K10: \citet{karakas10}
    \item V13: \citet{ventura+13}
    \item K16: \citet{KL16} \citet{karakas+18}
\end{itemize}
I use C11 as the fiducial AGB yield.
For an element X and star with mass $M$, the stellar yield $\tilde{y}$ is defined as the net production of X relative to $M$, or
\begin{equation}
    \tilde{y}_{\rm X} = \frac{M_{\rm X,\ ejected} - Z_{0, X} M_{\rm ejected}}{M}   
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm X, ejected}$  are the total ejected mass of the envelope and the element X, respectively.

Fig. \ref{fig:y_agb} compares the AGB C yield $\tilde{y}_\text{C}^\text{AGB}(M, Z)$ for each model I consider. 
The yields may be negative when the birth abundance is higher than the C abundance of the material ejected back to the ISM. 
Most models agree on the qualitative shape of the net fractional AGB C yield---% 
stars with mass $\sim 2--4 M_\odot$ have highest fractional C yields, with the mass of the peak increasing and overal yields decreasing with increasing $Z$.  High mass, high $Z$ stars destroy \isotope[12]{C} due to HBB.

In Fig. \ref{fig:agb-ssp}, we plot the total production of C by AGB stars as a function of SSP age.
Effectively, this is equivalent to integrating Eq. \ref{eq:imf-yield} down to $M_{to}$, the turnoff mass for time $t$. 
\footnote{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr,
from \citet{larson74}.}
As the mass range $2\lesssim M \lesssim 4$ is most important for C production, the characteristic delay time is $\sim 1$Gyr. In contrast, Fe from SNeIa (plotted as the dashed black line in Fig \ref{fig:agb-ssp}) has a longer characteristic delay time.

K10 and K16 weight C production more heavily towards massive stars resulting in a faster enrichment delay time, whereas the C11 and V13 models predict a slightly longer timescale with order 1Gyr, but little to no C is produced more than 2Gyr after a star formation event. This is in contrast to iron which is still produced by SNeIa up to 10Gyr after a star formation event. 
As shown in the right panel of Fig. \ref{fig:agb-ssp}, with increasing $Z$, C enrichment occurs earlier, and C destruction in low mass stars leads to a declining C abundance at late times.

\begin{figure}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[AGB C yields]{The net fractional AGB C yield ($\tilde{y}$) plotted as a function of initial stellar mass $M$. Metallicity is colorcoded according to the colorbar. I plot $\tilde{y}=0$ for reference. Each panel represents yields from one of four AGB models: C11 \citep{cristallo+11, cristallo+15}, K10 \citep{karakas10}, V13 \citep{ventura+13,ventura+14,ventura+18}, and K16 \citep{KL16, karakas+18}.
        }
        \label{fig:y_agb}
\end{figure}
\begin{figure}
    \includegraphics[scale=1]{y_agb_t2.pdf}

@@

Though per-star yieldds are necessary to compute AGB star enrichment rates in GCE models, IMF-averaged yields are useful in interpreting their predictions. For a yield $y$ from a star of mass M and initial metallicity $Z$, the IMF-averaged yield is given by 
More useful for GCE are Initial Mass Function (IMF) yields, where I sum up the
contributions of all stars which produce an element.
I define the IMF-integrated yield of X with:

\begin{equation} \label{eq:imf-yield}
    y_{\rm X}^{\rm proc}(Z) = 
    \int_{M_{\rm min}}^{M_{\rm max}} 
    \tilde{y}_{\rm X}^{\rm proc}(M, Z)
    \frac{dN}{dM}  \ dM
\end{equation}

where $\frac{dN}{dM}$ is the normalized IMF, $M_{\rm min}$ and $M_{\rm max}$ are the minimum and maximum mass of stars, which I take to be $0.08 M_{\sun}$ and $100 M_{\sun}$ respectively. For AGB stars, the yields are truncated above 8 $M_{\odot}$.

@@

In Fig. \ref{fig:y_cc}, I plot calculations of the IMF-integrated yields, using Eq. \ref{eq:imf-yield} (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function). On the right, I also show the CCSNe-[C/Mg] ratio, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
In other words, ${\rm [C/Mg]^{CC}}$ describes what [C/Mg] would be if CCSNe were the only process producing C.
Notice that even at $Z=Z_\odot$, there is a wide range ($\sim 1$ dex) in predicted $\Ycc$.

    \begin{itemize}
        \item Massive stars should dominate C yields, GCE models indicate similar results \citep[e.g.][]{prantzos+18,rybizki+17}.  
        \item Rotation in LC18 appears to increase metallicity dependence of C production (read this paper more carefully)
        \item NKT and LC (with rotation)agree in positive metallicity depenent yields
        \item [C/Mg]
        \item The fiducial value, where I choose the C yield from Sukhbold+16, may be above most predictions of C yields but is within the range of yield ratios
        \item the fiducial value would reach an equilibrium C yield of -0.1 without AGB stars
    \end{itemize}

@@

I define $M_{X}$ to be the mass of element X in the gas phase, $\dot{M}_\star$ to be the star fomation rate (in M$_\odot$ yr$^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ representing the strength of outflows. 
SSPs return a fraction $r$ of their birth mass back to the ISM, due to ejected stellar envelopes. $r\approx0.4$ for a \citet{kroupa01} IMF.
Given the IMF-weighted yield of Mg $y_{\rm Mg}$, the rate of change in the ISM mass of Mg $\dot{M}_{\rm Mg}$ is a simple sum of sources and sinks,
\begin{equation}
    \dot{M}_{\rm Mg} = \dot{M}_{\rm Mg,~new} - \dot{M}_{\rm Mg~remnants} - \dot{M}_{\rm Mg,~outflows},
\end{equation}
where the first term on the right-hand side describes CCSNe enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_X (1-r)\dot{M}_\star$.  And, the outflows deplete a mass $Z_X \eta\dot{M}_\star$. Substituting, 
\begin{equation}
    \dot{M}_{\rm Mg}= y_{X} \dot{M}_\star - (1 + \eta - r) Z_{X} \dot{M}_\star.
\end{equation}
Assuming an exponentially declining star formation history, $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, this equation can be solved analytically. 
To find equilibrium, set $\dot{M}_X=0$, and now
\begin{equation}\label{eq:z_eq}
    Z_{\rm Mg}^{\rm eq}(R) = \frac{y_{\rm Mg}}{1 + \eta(R) - r - \tau_\star / \tau_{sfh}}
\end{equation}
For constant star formation, $\tau_\star/\tau_{\rm sfh} \rightarrow 0$

While Equation \ref{eq:z_eq} can be applied to CCSNe C, the delayed time contribution of AGB C can complicate the result. Instead, using an effective or DTD-integrated yield, $\langle \Ycagb \rangle$
\begin{equation}
    \langle \Ycagb\rangle = \frac{\int_0^T \tilde{y}_{\rm C}^{\rm agb}(M, Z) \dot{M}_\star(T - t) \frac{dN}{dM} \frac{dM}{dt} dt  }{ \dot{M}_\star \int_0^T \frac{dN}{dM} \frac{dM}{dt} dt}
\end{equation}
So, for C, Equation \ref{eq:z_eq} becomes
\begin{equation}
    Z_{\rm C}^{\rm eq}(R) = \frac{\Ycc + \langle\Ycagb\rangle}{1 + \eta(R) - r - \tau_\star / \tau_{sfh}}
\end{equation}
Taking the ratio of $Z_{\rm C}$ and $Z_{\rm Mg}$, 
\begin{equation}\label{eq:z_co}
    \frac{Z_{\rm C}^{\rm eq}}{Z_{\rm Mg}^{\rm eq}} = \frac{\Ycc + \langle \Ycagb \rangle }{\Yoc}
\end{equation}

With these definitions, I can calculate the core collapse yields of C as a function of metallicity:
\begin{equation}
    y_\text{C}^\text{CC} =  y_\text{Mg}^\text{CC} \frac{Z_\text{C,~eq}}{Z_\text{Mg,~eq}} - \langle y_c^{agb} \rangle
\end{equation}

Rewriting in terms of the observed abundance ratio and dividing by
$y_\text{Mg}^\text{CC}$. So, given the outflows or CCSNe O yields, the
C AGB yield, and the \caah mean abundance track, I can estimate the CCSNe C yield. Written as a relative yield:
\begin{equation}
    \frac{\Ycc}{\Yoc} = \frac{Z_{\text{C},~\sun}}{Z_{\text{Mg},~\sun}} 10^{[C/Mg]} - \frac{\langle \Ycagb \rangle}{\Yoc}
\end{equation}

Fitting this relationship to data (figure \ref{fig:analytic}).

\begin{figure}
    \centering
    \includegraphics[]{analytic.pdf}
    \caption[Reverse fit yields]{Inferred massive star C yields as a function of metallicity. I assume equilibrium and $3\times {\rm C11}$ AGB yields (orange curve, see discussion in Section \ref{sec:equilibrium}). The binned medians and median absolute devation (divided by $\sqrt{N}$) are blue points.
    }
\end{figure}

@@

@@

@@

Even changing the AGB model does not affect the conclusion significantly.

>>>>

\begin{titlepage}
   \begin{center}
       \vspace*{5\baselineskip}
       \textbf{The Galactic Chemical Evolution of Carbon} \\
       Implications for Stellar Nucleosynthesis\\
       \vspace*{3\baselineskip}
        Undergraduate Research Thesis\\
       \vspace*{3\baselineskip}
    Presented in partial fulfillment of the requirements for graduation \textit{with research distinction in Astronomy and Astrophysics} in the College of Arts and Sciences of The Ohio State University \\
       \vspace*{3\baselineskip}
        by \\
       \vspace*{3\baselineskip}
       {Daniel Alexander Boyea}\\
       \vspace*{3\baselineskip}
       The Ohio State University\\
       April 2023\\
       \vspace*{3\baselineskip}
       Project Advisor: David Weinberg, Department of Astronomy
       % Rolando Vald\'es Aguilar\\
       \vfill
   \end{center}
\end{titlepage}

@@

Here, I aim to understand C enrichment---where it is produced and how its abundances evolve over time. Why study C? C is unique nucleosynthetically, one of few light elements (along with N) produced in asymptotic giant branch (AGB) stars \citep[e.g.][]{jennifer19, KL14}. C and N are well studied elements as they are easy to observe in stellar spectra, even at the lowest metallicities \citep{FN15}(?). Additionally, C and N are used as age indicators in red giant branch stars \citep[e.g.][]{martig16, MG15, fiorenzo+21}.

Gas-phase [C/O] ratios---observed in very low metallicity, high redshift damped Lyman-alpha systems (DLAs)---decrease with increasing [O/H] \citep{cooke+17}. Then, [C/O] increases above $\rm [O/H] \approx 1$ (\citealt{berg+19,nissen+14,fabbian+09}; see discussion in \S\ref{sec:gas}).
While we know C is produced in AGB stars and core collapse supernovae (CCSNe), we still have limited understanding of the magnitude and metallicity-dependence of each process.

Despite their central role in GCE models, nucleosynthetic yields predicted by stellar evolution models are rife with uncertainties. Poorly understood processes, including mass loss, opacity, nuclear reaction rates, rotational mixing, and convection/convective boundaries each shape the production rates of atomic nuclei in stellar interios \citep{KL14,ventura+13, LC18}.
Additionally, changes in massive star's explodability can introduce a factor of $\sim2$ change in population averaged-yields \citep{emily+21}.

GCE models typically use unmodified nucleosythetic predictions and attempt to create models matching observations by varying the star formation history (SFH) and other evolutionary parameters. I choose to introduce C yields as a free parameter, determining which yield prescriptions reproduce Galactic abundance trends.
\cite{james+22} examined similar GCE models of N, an elements whose production is closely related to C. \cite{james+22} found trends in N and O abundances can be explained by the metallicity dependence of relative N and O yields. Here, I extend their models to C and show that similar constraints on C and O relative yields can be obtained from observed abundance ratios. Additionally, I assess which yield prescriptions reproduce Galactic abundance trends, while assessing the impact of SFH and GCE model assumptions.

Accurate determination of stellar birth abundances for C and N is challenging. When a star enters the red giant branch (RGB), material from the CNO-processed core is mixed with the envelope in first dredge up (FDU), enhancing N and depleting C \citep{fiorenzo+21,KL14}. Measurements of these evolved stars'  atmospheres will no longer reflect birth abundances.  Additionally, C is faint in HII regions and determining gas-phase abundances is challenging \citep{skillman+20}.

To this end, I use a sample of subgiant stars from APOGEE \citep{apogee17}. According to stellar evolution models (CITES), these stars have not yet experienced FDU. Their phosphoric abundances should thus represent their birth composition.  In Fig. \ref{fig:subgiants}, I plot this sample, selected by criteria in \citet[see Appendix \ref{sec:jack}]{jack_subgiant}. These emperical resolults contain information on the relative C and Mg yields, which I use later in this work.

@@

Theoretical models of stellar nucleosynthesis provide the starting point of this investigation. I focus on three primary nucleosynthetic pathways: Asymptotic giant branch (AGB) stars, core collapse supernovae (CCSNe), and type Ia supernovae (SNe Ia). Each process has unique timescales and yields, traceable through the tools of GCE. While C is only produced in AGB stars and CCSNe, comparing C to Fe, which is produced in SNe Ia, adds a valuable constraint on the delay-time distribution (DTD) of C enrichment.

When a single stellar population (SSP) forms, CCSNe are the first enrichment events. CCSNe provide light elements such as C, O, and Mg and heavier elements such as Fe and beyond. Metallicity-independent yields from CCSNe are the only statistically significant source of O and Mg ($\alpha$ elements). Next, low mass stars begin to reach the end of their lives. The dying breaths of AGB stars are known to be important sources of C, N, and neutron capture elements.  Finally, white dwarfs explode, releasing Fe and other iron-peak elements.

For an element X and star with mass $M$, the stellar yield $\tilde{y}$ is defined as the net production of X relative to $M$, or
\begin{equation}
    \tilde{y}_{\rm X} = \frac{M_{\rm X,\ ejected} - Z_{0, X} M_{\rm ejected}}{M}   
\end{equation}
where $M_{\rm ejected}$ and $M_{\rm X, ejected}$  are the total ejected mass of the envelope and the element X, respectively.

Though per-star yields are necessary to compute AGB star enrichment rates in GCE models, IMF-averaged\footnote{IMF: Initial Mass Function} yields are useful in interpreting their predictions. For a yield $y$ from a star of mass M and initial metallicity $Z$, the IMF-averaged yield is given by 
\begin{equation} \label{eq:imf-yield}
    y_{\rm X}^{\rm proc}(Z) = 
    \int_{M_{\rm min}}^{M_{\rm max}} 
    \tilde{y}_{\rm X}^{\rm proc}(M, Z)
    \frac{dN}{dM}  \ dM
\end{equation}
where ${dN}/{dM}$ is the normalized IMF, $M_{\rm min}$ and $M_{\rm max}$ are the minimum and maximum mass of stars, which I take to be $0.08 M_{\sun}$ and $100 M_{\sun}$ respectively. For AGB stars, the yields are truncated above 8 $M_{\odot}$. 
To calculate the IMF-weighted net yields of C, I use the Versatile Integrator for Chemical Evolution code (\VICE\footnote{\VICE~is available at \url{https://github.com/giganano/VICE}}).

In the interest of focusing on carbon yields, I hold yields of other elements fixed. 
I adopt yields from \citet{james+21, james+22}.
In Table \ref{tab:fiducial_mod}, I tabulate yield values, in units of a stellar population's birth mass.
Also following \citet{james+21, james+22}, I take the SNeIa delay time distribution to be a
$t^{-1.1}$ law suggested by the observations of \citet{maoz+12}.

@@

An AGB star is a low mass ($\lesssim 8 M_{\sun}$) star during its final phase of evolution.  In an AGB star, two competing processes determine the outcome of C production: third dredge up (TDU) and hot bottom burning (HBB).  TDU accompanies thermal pulsations in AGB stars. During a TDU event, material from the CO core is mixed with material exposed to partial He-shell burning, increasing C and O abundances in the envelope \citep{KL14}.
HBB is the activation of proton capture reactions, and the CNO cycle%
\footnote{The CNO (I) cycle is
\isotope[12]{C}(p, $\gamma$)
\isotope[13]{N}($\beta^+ \nu_e$)
\isotope[13]{C}(p, $\gamma$)
\isotope[14]{N}(p, $\gamma$)
\isotope[15]{O}($\beta^{+}\nu_e$)
\isotope[15]{N}(p, \alp)
\isotope[12]{C}.
 Other minor branches of the CNO cycle are not relevant here.
 \citep{solar-fusion}.
}
at the base of the convective envelope. Because the $^{14}$N(p, $\alpha$) proton capture is the slowest component of the CNO cycle \citep{solar-fusion}, the CNO cycle converts nearly all \isotope[12]{C} into \isotope[14]{N}.
As a result, when both TDU and HBB occur, \isotope[12]{C} yields are lowered (see discussion in \citealt{james+22} and \citealt{ventura+13}).

In this work, I explore four different sets of AGB star yield tables from literature, providing well sampled grids in metallicity and mass for use in chemical evolution models. 
\begin{itemize}
    \item C11: \citet{cristallo+11, cristallo+15}
    \item K10: \citet{karakas10}
    \item V13: \citet{ventura+13,ventura+14,ventura+18}
    \item K16: \citet{KL16} \citet{karakas+18}
\end{itemize}
I use C11 as the fiducial AGB yield.

Fig. \ref{fig:y_agb} compares the AGB C yield $\tilde{y}_\text{C}^\text{AGB}(M, Z)$ for each model I consider. 
The yields may be negative when the birth abundance is higher than the C abundance of the material ejected back to the interstellar medium (ISM). 
Most models agree on the qualitative shape of the net fractional AGB C yield---% 
stars with mass $\sim 2--4 M_\odot$ have highest fractional C yields, with the mass of the peak increasing and overal yields decreasing with increasing $Z$.  High mass, high $Z$ stars destroy \isotope[12]{C} due to HBB.

In Fig. \ref{fig:agb-ssp}, we plot the total production of C by AGB stars as a function of SSP age.
Effectively, this is equivalent to integrating Eq. \ref{eq:imf-yield} down to $M_{to}$, the turnoff mass for time $t$%
\footnote{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr,
from \citet{larson74}.}.
As the mass range $2\lesssim M \lesssim 4$ is most important for C production, the characteristic delay time is $\sim 1$Gyr. In contrast, Fe from SNeIa (plotted as the dashed black line in Fig \ref{fig:agb-ssp}) has a longer characteristic delay time.

K10 and K16 weight C production more heavily towards massive stars resulting in a faster enrichment delay time, whereas the C11 and V13 models predict a slightly longer timescale with order 1 Gyr, but little to no C is produced more than 2 Gyr after a star formation event. This is in contrast to iron which is still produced by SNeIa up to 10 Gyr after a star formation event. 
As shown in the right panel of Fig. \ref{fig:agb-ssp}, with increasing $Z$, C enrichment occurs earlier, and C destruction in low mass stars leads to a declining C abundance at late times.

\begin{figure}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[AGB C yields]{The net fractional AGB C yield ($\tilde{y}$) plotted as a function of initial stellar mass $M$. Metallicity is colorcoded according to the colorbar. I plot $\tilde{y}=0$ for reference. Each panel represents yields from one of four AGB models: C11, K10, V13, K16 (see \S \ref{sec:agb}). }
        \label{fig:y_agb}
\end{figure}
\begin{figure}
    \includegraphics[scale=1]{y_agb_t2.pdf}

@@

@@

In Fig. \ref{fig:y_cc}, I plot calculations of the IMF-integrated yields, using Eq. \ref{eq:imf-yield} (computed using \VICE's \texttt{vice.yields.ccsne.fractional} function). 
CCSNe models predict a wide range of C yields, spanning almost 1 dex. 
Both the \citet{NKT13} and \cite{LC18} models show positive metallicity dependence. 
The \cite{LC18} models also include rotation, showing that variations in the rotational velocity of the star can dramatically increase C production and the metallicity dependence of $\Ycc$. Rotation induces more mixing allowing the CO core to grow larger. Our fiducial model's steep metallicity dependence near $Z\approx Z_\odot$ could be explained by rotation.

I show the C11 model for comparison in Fig. \ref{fig:y_cc}. Especially at $Z\approx Z_\odot$, most CCSNe models would dominate C production. I will later show that empirically, CCSNe also dominate C production.

On the bottom of Fig. \ref{fig:y_cc}, I also show the CCSNe-[C/Mg] ratio, defined by
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/Mg]^{CC}} = \log_{10}\left( \frac{\Ycc}{\Yoc}\right) - \log_{10} \left( \frac{Z_{{\rm C},\ \sun }}{Z_{{\rm Mg},\ \sun }} \right).
\end{equation}
In other words, ${\rm [C/Mg]^{CC}}$ describes what [C/Mg] would be if CCSNe were the only process producing C.
Once again, different CCSNe models span a large range in [C/Mg]. 
I chose to instead parameterize $\Ycc$ to enable agreement with observations, as most CCSNe models fail to achieve near-solar [C/Mg].

Assumptions about the explodability landscape affect C and Mg production. Increasing the fraction of stars which explode increases $\Ycc$, as stars which directly collapse do not contribute to explosive yields \citep{emily+21}. However, C is relatively unaffected by the black-hole landscape, as very massive stars contribute carbon through enriched winds. Since Mg is formed deeper in the core of massive stars, the Mg yield drops much more steeply, so models where few stars explode (S16/W18) result in much higher [C/Mg].

CCSNe models also struggle to reach solar [O/Mg], either due to overproduction of O or underproduction of Mg \citep{emily+21}. Here, I assume ${\rm [O/Mg]} = 1$, which is not compatible with CCSNe models but is consistent with observations \cite{Weinberg+19, Weinberg+22}.

@@

I define $M_{X}$ to be the mass of element X in the gas phase, $\dot{M}_\star$ to be the star fomation rate (in M$_\odot$ yr$^{-1}$), and $\eta$ to be the mass loading factor $\eta\equiv\dot{M}_{\rm outflow}/\dot{M}_\star$ representing the strength of outflows. 
SSPs return a fraction $r$ of their birth mass back to the ISM, due to ejected stellar envelopes. $r\approx0.4$ for a \citet{kroupa01} IMF.
Given the IMF-weighted yield of Mg $y_{\rm Mg}$, the rate of change in the ISM mass of Mg $\dot{M}_{\rm Mg}$ is a simple sum of sources and sinks,
\begin{equation}
    \dot{M}_{\rm Mg} = \dot{M}_{\rm Mg,~new} - \dot{M}_{\rm Mg~remnants} - \dot{M}_{\rm Mg,~outflows},
\end{equation}
where the first term on the right-hand side describes CCSNe enrichment. 
In terms of the return mass fraction of stars $r$, the mass lost to remnants is $Z_X (1-r)\dot{M}_\star$.  And, the outflows deplete a mass $Z_X \eta\dot{M}_\star$. Substituting om fpr $\eta$ and $r$,  
\begin{equation}
    \dot{M}_{\rm Mg}= y_{X} \dot{M}_\star - (1 + \eta - r) Z_{X} \dot{M}_\star.
\end{equation}
Assuming an exponentially declining star formation history $\dot{M}_\star \propto e^{-t/\tau_{\rm sfh}}$, the equilivrium abundance is derived analytically by setting $\dot{Z}_{Mg}=0$.
\begin{equation}\label{eq:z_eq}
    Z_{\rm Mg}^{\rm eq}(R) = \frac{y_{\rm Mg}}{1 + \eta(R) - r - \tau_\star / \tau_{sfh}}
\end{equation}
In the special case of constant star formation, $\tau_{\rm sfh}\to\infty$, simplifying the denominator to $1+\eta-r$.

While Equation \ref{eq:z_eq} can be applied to CCSNe C, the delayed nature of AGB C can complicates thes expression. Instead, using a DTD-integrated yield, $\langle \Ycagb \rangle$ enrichment is
\begin{equation}
    \langle \Ycagb\rangle = \frac{\int_0^T \tilde{y}_{\rm C}^{\rm agb}(M, Z) \dot{M}_\star(T - t) \frac{dN}{dM} \frac{dM}{dt} dt  }{ \dot{M}_\star \int_0^T \frac{dN}{dM} \frac{dM}{dt} dt}
\end{equation}
Eq. \ref{eq:z_eq} becomes
\begin{equation}
    Z_{\rm C}^{\rm eq}(R) = \frac{\Ycc + \langle\Ycagb\rangle}{1 + \eta(R) - r - \tau_\star / \tau_{sfh}}
\end{equation}
Taking the ratio of $Z_{\rm C}$ and $Z_{\rm Mg}$, 
\begin{equation}\label{eq:z_co}
    \frac{Z_{\rm C}^{\rm eq}}{Z_{\rm Mg}^{\rm eq}} = \frac{\Ycc + \langle \Ycagb \rangle }{\Yoc}.
\end{equation}

Analogous to \cite{james+22} arguments about N, the trends in abundance ratios are set by yield ratios in these GCE models. The effect of other GCE parameters (most importaintly $\eta$) cancels. As a consequence, yield ratios should establish abundance ratio trends in models which assume a different normalization of element yields and mass-loading (see discussion below).
This argument can also be inverted to infer  yields from abundance ratio trends. To the extent that observed C and Mg trends reflect the equilibrium abundances in different Galatic regions, we can infer the CCSNe (is this plural><?) yield given an assumed AGB star yield (or vice versa). Inffering $\Ycc$ from $\Ycagb$, 
\begin{equation}
    y_\text{C}^\text{CC} =  y_\text{Mg}^\text{CC} \frac{Z_\text{C,~eq}}{Z_\text{Mg,~eq}} - \langle y_c^{agb} \rangle
\end{equation}
Rewriting this expression as a relative yield of C and Mg,
\begin{equation}
    \frac{\Ycc}{\Yoc} = \frac{Z_{\text{C},~\odot}}{Z_{\text{Mg},~\odot}} 10^{\rm [C/Mg]} - \frac{\langle \Ycagb \rangle}{\Yoc}.
\end{equation}

@@

If $\Ycc$ gets us to ${\rm [C/Mg] } \approx -1$, then $\Yct = 0.005$, but if 0, $y\approx0.006$. 
Section 2.2, additional AGB yield 1part per thousand?, Motivate AGB yield ...

@@

Fitting this relationship to data (figure \ref{fig:analytic}).
I find that a linear dependence on metallicity,

\begin{equation}
    \frac{\Ycc}{\Yoc} = y_{\rm}
\end{equation}

\begin{figure}
    \centering
    \includegraphics[]{analytic.pdf}
    \caption[Reverse fit yields]{Inferred massive star C yields as a function of metallicity. I assume equilibrium and $3\times {\rm C11}$ AGB yields (orange curve, see discussion in Section \ref{sec:equilibrium}). The binned medians and median absolute devation (divided by $\sqrt{N}$) are blue points.
    }
    \label{fig:analytic}
\end{figure}

@@
